---
title: "SSJ-DELTA-CU report figures"
output:
  html_document: default
  html_notebook: default
---

```{r "setup", include=FALSE}
require("knitr")
opts_knit$set(root.dir = "..") # set working directory to be ssj-graph-r directory
knitr::opts_chunk$set(error = TRUE)
```


# About
This Rmarkdown document is for generating the plots for the report. The chunks should be run individually.

```{r output_settings, echo=FALSE, include=FALSE}
source("scripts/helper_functions.R") # load all the helper functions
source("scripts/figures.R") # load all the functions that generate the graphs

# data to use for generating the plots
dsa_legal_data <- readRDS("data/dsa_legal/dsa_legal_5methods_20170517.rds")
fieldstations_data <- readRDS("data/fieldstations/wy2016/fieldstations_wy2016_20170517.rds")


# list of crops to generate
crops_2_gen_all <- crop_list() 
crops_2_gen <- c("Alfalfa", "Almonds", "Corn", "Pasture", "Potatoes", "Rice", "Tomatoes", "Vineyards")

# list of months of the water year (Oct-> sept)
month_2_gen <- 'JUL' #month_list()

# lsit of methods to use for the plots
methods_2_gen <- methods_names
methods_2_gen <- methods_2_gen[!methods_2_gen %in% c("CalSIMETAW", "DETAW")]

# region of interest (usually dsa or legal)
aoi <- "dsa"

# water years
water_years <- c(2015, 2016)

# footer 
footer <- paste('Updated ', Sys.Date())

# creates a figure output folder if one does not already exist
fignum_folder <- function(fig_number){
  mainDir <- getwd() # gets the current working directory 
  subDir <-paste("figs/fig", fig_number, sep="") # construct path to subfolder
  ifelse(!dir.exists(file.path(mainDir, subDir)), 
         dir.create(file.path(mainDir, subDir)), FALSE) # creates subDir if it doesn't exist
  subDir_path <- file.path(mainDir, subDir) # returns the path to the new subDir
  return(subDir_path)
}


```

# Field

## Field vs Remote methods facet plot

```{r field_v_remote_facet, echo=FALSE, fig.width = 11, fig.height = 5}

# figure number
fig_num <- 5

# make an ouput folder for the figure
output_folder <- fignum_folder(fig_num)


# facet version
p <- lm_fieldxremote_facet_crop(fieldstations_data)

# add footer
p <- p + labs(caption=footer)


# build name for export
base <- paste("Field_vs_Model_facet_crops", ".png", sep='')
name <- paste(output_folder, base, sep="/")

# save
ggsave(name, p, width=11, height=5, units="in")

# add plot to rmarkdown output
plot(p)

######################################################################
# 
# p1 <- lm_fieldxremote_crop(fieldstations_data, "pasture")
# p1
# 
# base <- paste("Field_vs_Model_pasture", ".png", sep='')
# name <- paste(file.path(mainDir, subDir), base, sep="/")
# 
# # save
# ggsave(name, p1, width=5, height=5, units="in")
# print(name)

```

# Delta

## Total Acre-feet Figures + csv files for Water Year 2015 + 2016

```{r TAF_WY, echo=FALSE, fig.width = 8, fig.height = 4}
source("scripts/tables.R") # load in the table file to calc acre-ft totals to export to csv.
source("scripts/figures.R")
# figure number
fig_num <- 6

# make an ouput folder for the figure
output_folder <- fignum_folder(fig_num)


# loop through both water years and save the figure
for(y in 1:length(water_years)){
  wy <- water_years[y]
  
  p <- sum_TAFxModel_dsa_legal(dsa_legal_data, wy)
  
  # add footer
  p <- p + labs(caption=footer)
  
  
  # build name for export
  base <- paste("TAF_wy", wy, "_dsa_legal", ".png", sep='')
  name <- paste(output_folder, base, sep="/")

  # save
  ggsave(name, p, width=8, height=4, units="in")
  
  # add plot to rmarkdown output
  plot(p)
  
  # csv files
  
  # let's also generate a table with this data and save as csv
  t <- table_taf_wy_regions(dsa_legal_data, wy)
  
  # build name for export
  base <- paste("TAF_wy", wy, "_dsa_legal", ".csv", sep='')
  name <- paste(file.path(mainDir, subDir), base, sep="/")
  
  write.csv(t, name)
  }

```

## Lineplot Monthly ET by Model for a single croptype for two water years

```{r lineplot_ET, echo=FALSE, fig.width = 7, fig.height = 4}
# batch outputs monthly et + ref Eto for both water years on one plot
fig_num <- 11

# make an ouput folder for the figure
output_folder <- fignum_folder(fig_num)

# loop through the crops
for(i in 1:length(crops_2_gen)){
  crop<-crops_2_gen[i]
  cropid <- lookup_cropid(crop)
  
  #plot
  p <- line_ETxMonths_2wateryears(dsa_legal_data, cropid,  aoi)
  
  # add footer
  p <- p + labs(caption=footer)
  
  # build ouput name
  crop <- gsub(" ", "", crop)
  crop <- gsub("/", "", crop)
  base <- paste(cropid, crop, aoi, sep="-")
  base.ext <- paste(base, ".png", sep="")
  name <- paste(output_folder, base.ext, sep="/")
  
  # save
  ggsave(name, p, width=7, height=4, units="in")
  
  # add plot to rmarkdown output
  plot(p)
}
```

## Cumulative Lineplot Monthly ET by Model for a single croptype for two water years

```{r lineplot_cumET, echo=FALSE, fig.width = 7, fig.height = 4}
# batch outputs monthly et + ref Eto for both water years on one plot
fig_num <- 12

# make an ouput folder for the figure
output_folder <- fignum_folder(fig_num)

# loop through the crops
for(i in 1:length(crops_2_gen)){
  crop<-crops_2_gen[i]
  cropid <- lookup_cropid(crop)
  
  #plot
  p <- line_ETxMonths_cumulative(dsa_legal_data, cropid,  aoi)
  
  # add footer
  p <- p + labs(caption=footer)
  
  # build ouput name
  crop <- gsub(" ", "", crop)
  crop <- gsub("/", "", crop)
  base <- paste(cropid, crop, aoi, sep="-")
  base.ext <- paste(base, ".png", sep="")
  name <- paste(output_folder, base.ext, sep="/")
  
  # save
  ggsave(name, p, width=7, height=4, units="in")
  
  # add plot to rmarkdown output
  plot(p)
}
```


## Boxplot whisker_ETxModel_per_crop_month

```{r whisker_ETxModel_per_crop_month_2yrs, echo=FALSE, fig.width = 7, fig.height = 4}
# batch outputs monthly et + ref Eto for both water years on one plot
fig_num <- 13

# make an ouput folder for the figure
output_folder <- fignum_folder(fig_num)


# loop through the crops
for(i in 1:length(crops_2_gen)){
  crop<-crops_2_gen[i]
  cropid <- lookup_cropid(crop)

  # loop through the months
  for(m in 1:length(month_2_gen)){
    month <- month_2_gen[m]

    #plot
    p <- whisker_ETxModel_per_crop_month_2yrs(dsa_legal_data, aoi, month, cropid)
    
    # add footer
    p <- p + labs(caption=footer)

    # build ouput name
    crop <- gsub(" ", "", crop)
    crop <- gsub("/", "", crop)
    base <- paste(cropid, crop, month, aoi, sep="-")
    base.ext <- paste(base, ".png", sep="")
    name <- paste(output_folder, base.ext, sep="/")
    
    # save
    ggsave(name, p, width=7, height=4, units="in")
    
    # add plot to rmarkdown output
    plot(p)
    
  }
}


```


## Boxplot whisker_ETxMonths_per_crop_method

```{r whisker_ETxMonths_per_crop_method, echo=FALSE, fig.width = 7, fig.height = 4}
# batch outputs monthly et + ref Eto for both water years on one plot
fig_num <- 14

# make an ouput folder for the figure
output_folder <- fignum_folder(fig_num)

# loop through all crops types
for(i in 1:length(crops_2_gen)){
  crop<-crops_2_gen[i]
  cropid <- lookup_cropid(crop)
  
  # loop through the months
  for(m in 1:length(methods_2_gen)){
    mod <- methods_2_gen[m]
    
    # plot
    p <- whisker_ETxMonths_per_crop_method_2yrs(dsa_legal_data, aoi, mod, cropid)
    
    # add footer
    p <- p + labs(caption=footer)
    
    # build ouput name
    crop <- gsub(" ", "", crop)
    crop <- gsub("/", "", crop)
    m_lower <- lookup_model_name(mod) # get lowercase name
    base <- paste(cropid, crop, m_lower, aoi, sep="-")
    base.ext <- paste(base, ".png", sep="")
    name <- paste(output_folder, base.ext, sep="/")
    
    # save
    ggsave(name, p, width=7, height=4, units="in")

    # add plot to rmarkdown output
    plot(p)
  }
}

```


## Grid coefficient variation

```{r coefficient_variation_grid, echo=FALSE, fig.width = 6, fig.height = 5}
# coeffient of variation grid
fig_num <- 24

# make an ouput folder for the figure
output_folder <- fignum_folder(fig_num)

# loop through both water years and save the figure
for(y in 1:length(water_years)){
  wy <- water_years[y]
  
  p <- coefficient_variation_grid(dsa_legal_data, wy)
  
  # add footer
  p <- p + labs(caption=footer)
  
  # build name for export
  base <- paste("CV_grid_", wy, ".png", sep='')
  name <- paste(output_folder, base, sep="/")
  
  # save
  ggsave(name, p, width=6, height=5, units="in")
  
  # add plot to rmarkdown output
  plot(p)
}

```