---
title: "SSJ-DELTA-CU report figures"
output:
  html_document: default
---

# About
This Rmarkdown document (`figs/figures.Rmd`) is used for generating the figures for the SSJ-Delta-CU report. The chunks can be run individually but make sure to run the `setup` and `project_settings` chunk first to ensure that all the libraries and data are loaded. 

```{r "setup", include=FALSE}
# knitr settings
require("knitr")
opts_knit$set(root.dir = "..") # set working directory to be ssj-graph-r directory
knitr::opts_chunk$set(error = TRUE)

```

```{r project_settings, echo=FALSE, include=FALSE}
# source scripts for plots and helper functions
source("scripts/helper_functions.R") # load all the helper functions
source("scripts/figures.R") # load all the functions that generate the graphs
source("scripts/tables.R") # load in the table file to calc acre-ft totals to export to csv.

# data sources
dsa_legal_data <- readRDS("data/dsa_legal/dsa_legal.rds")
dsa_etrf <- readRDS("data/dsa_etrf/dsa_etrf.rds")
delta_regions <- readRDS("data/delta_regions/delta_regions.rds")
fieldstations_data <- readRDS("data/fieldstations/wy2016/fieldstations_wy2016.rds")
fieldstations_data_2015 <- readRDS("data/fieldstations/wy2015/fieldstations_wy2015.rds")
subregions <- readRDS("data/subregions/subregions.rds")
eto <- readRDS('data/eto/eto.rds')

# vector containing the water years
water_years <- c(2015, 2016)

# footer 
footer <- paste('Updated ', Sys.Date())

```

# Delta Level Analysis

## Total Acre-feet for Legal Delta and Delta Service Area

```{r TAF_WY, echo=FALSE, fig.width = 8, fig.height = 4}

# loop through both water years 
for(y in 1:length(water_years)){ 
  
  wy <- water_years[y]
  
  # create the sum_TAF plot for the selected water year
  p <- sum_TAFxModel_dsa_legal(dsa_legal_data, wy)
  
  # add footer (commented out to avoid overwriting the existing caption)
  #p <- p + labs(caption=footer)
  
  # let's also generate a table with this data.
  t <- table_taf_wy_regions(dsa_legal_data, wy)
  
  # add plot and table to rmarkdown output
  plot(p)
  print(t)
  }

```

## Lineplot Monthly ET by Model for a single crop for two water years

```{r lineplot_ET, echo=FALSE, fig.width = 7, fig.height = 4}

# list of crops to generate
crops_2_gen <- c("Alfalfa", "Almonds", "Corn", "Pasture", "Potatoes", "Rice", "Tomatoes", "Vineyards")

# loop through the crops
for(i in 1:length(crops_2_gen)){
  
  # get the id from the crop name
  cropid <- lookup_cropid(crops_2_gen[i])
  
  # generate plot for the crop for the delta service area
  p <- line_ETxMonths_2wateryears(dsa_legal_data, cropid,  "dsa")
  
  # add footer with the date
  p <- p + labs(caption=footer)
  
  # add plot to rmarkdown output
  plot(p)
}
```

## Cumulative Lineplot Monthly ET by Model for a single crop for two water years

```{r lineplot_cumET, echo=FALSE, fig.width = 7, fig.height = 4}

# list of crops to generate
crops_2_gen <- c("Alfalfa", "Almonds", "Corn", "Pasture", "Potatoes", "Rice", "Tomatoes", "Vineyards")

# loop through the crops
for(i in 1:length(crops_2_gen)){

  # get the id from the crop name
  cropid <- lookup_cropid(crops_2_gen[i])
  
  #plot
  p <- line_ETxMonths_cumulative(dsa_legal_data, cropid,  "dsa")
  
  # add footer
  p <- p + labs(caption=footer)
  
  plot(p)
}
```



## Boxplot whisker_ETxModel_per_crop_month

```{r whisker_ETxModel_per_crop_month_2yrs, echo=FALSE, fig.width = 7, fig.height = 4}

# list of crops to generate
crops_2_gen <- c("Alfalfa", "Almonds", "Corn", "Pasture", "Potatoes", "Rice", "Tomatoes", "Vineyards")

# list of the months in the water year to generate (Oct-> sept)
month_2_gen <- c('JUL') #month_list()


# loop through the crops
for(i in 1:length(crops_2_gen)){

  # get the id from the crop name
  cropid <- lookup_cropid(crops_2_gen[i])

  # loop through the months
  for(m in 1:length(month_2_gen)){
    month <- month_2_gen[m]

    #plot
    p <- whisker_ETxModel_per_crop_month_2yrs(dsa_legal_data, 'dsa', month, cropid)

    # add footer
    p <- p + labs(caption=footer)

    # add plot to rmarkdown output
    plot(p)
    
  }
}


```


## Grid coefficient variation

```{r coefficient_variation_grid, echo=FALSE, fig.width = 6, fig.height = 5}
# coeffient of variation grid

# loop through both water years and save the figure
for(y in 1:length(water_years)){
  wy <- water_years[y]
  
  p <- coefficient_variation_grid(dsa_legal_data, wy)
  
  # add footer
  p <- p + labs(caption=footer)

  # add plot to rmarkdown output
  plot(p)
}

```

## Grid absolute variation

```{r absolute_variation_grid, echo=FALSE, fig.width = 6, fig.height = 5}
# absolute variation grid

# loop through both water years and save the figure
for(y in 1:length(water_years)){
  wy <- water_years[y]
  
  p <- absolute_variation_grid(dsa_legal_data, wy)
  
  # add footer
  p <- p + labs(caption=footer)

  # add plot to rmarkdown output
  plot(p)
}

```

## Percent Total by the top crop types

Note: this doesn't include wet herb or semi-ag landuse classes.


```{r landuse_percent, echo=FALSE, fig.width = 9, fig.height = 4}

# filter out wet herb and landuse
dsa_legal_data_wo_wetsemi <- dsa_legal_data %>% filter(!level_2 %in% c(913, 2008))

for(y in 1:length(water_years)){
  wy <- water_years[y]
  p <- barchart_TAFxModel_PERCENT_wy_by_topcrops(dsa_legal_data_wo_wetsemi, wy, 'dsa')
  
  # add footer
  p <- p + labs(caption=footer)

  plot(p)
}

```

## Total Acre feet by month and crop type

Note: this doesn't include wet herb or semi-ag landuse classes.

The abberivation for the model groups are: "calsimetaw"="C", "detaw"="D", "disalexi"="D", "itrc"="I", "sims"="S", "ucdmetric"="M", "ucdpt"="P".

```{r crop breakdown_facet, echo=FALSE, fig.width = 11, fig.height = 4}

# filter out wet herb and landuse
dsa_legal_data_wo_wetsemi <- dsa_legal_data %>% filter(!level_2 %in% c(913, 2008))

for(y in 1:length(water_years)){
  wy <- water_years[y]
  p <- barchart_TAFxModel_sum_monthsfacet_by_topcrop(dsa_legal_data_wo_wetsemi, wy, 'dsa')
  
  # add footer
  p <- p + labs(caption=footer)

  plot(p)
}

```

```{r, fig.width = 11, fig.height = 4}
data <- dsa_legal_data
water_year <- 2015
aoi <- 'dsa'

  top_crops <- c("Alfalfa", "Corn",  "Pasture", "Tomatoes", "Vineyards")

  af_crops_top_only <- data %>%
    filter_no_eto() %>%
    filter(wateryear == water_year) %>%
    filter(region == aoi) %>%
    filter(include == "yes") %>%
    filter(cropname %in% top_crops) %>%
    dplyr::group_by(model, cropname, month) %>%
    dplyr::summarise(sum_af=sum(crop_acft))

  af_crops_others <- data %>%
    filter_no_eto() %>%
    filter(wateryear == water_year) %>%
    filter(region == aoi) %>%
    filter(include == "yes") %>%
    filter(!cropname %in% top_crops) %>%
    mutate(cropname = "Other") %>%
    dplyr::group_by(model, cropname, month) %>%
    dplyr::summarise(sum_af=sum(crop_acft))

  #  merge af_crops_top_only and af_crops_others
  af_crops_w_others <- dplyr::union(af_crops_top_only, af_crops_others)

  # modify axis units (acre-ft -> thousands acre-feet)
  axis_units <-function(x){
    x/1000
  }
  
  methods_first_letter_list <- c("calsimetaw"="C", "detaw"="D", "disalexi"="D", "itrc"="I", "sims"="S", "ucdmetric"="M", "ucdpt"="P")
  
   months_first_letter_list <- c("O", "N", "D", "J", "F", "M", "A", "M", "J", "J", "A", "S")
  
   facet_labels <- function(variable,value){
    return(methods_named_list[value])
  }
   

  # change order of months to follow water year Oct -> Sept
  af_crops_w_others$month <- factor(af_crops_w_others$month, levels=c("OCT", "NOV", "DEC", "JAN", "FEB", "MAR", "APR", "MAY", "JUN", "JUL", "AUG", "SEP"))

  levels(af_crops_w_others$model) <- c("CalSIMETAW", "DETAW", "DisALEXI", "ITRC", "SIMS", "UCD-METRIC", "UCD-PT")
  
  p <- ggplot(af_crops_w_others, aes(x=month, y=sum_af, fill=cropname))+geom_bar(stat = "identity") +
    facet_grid(. ~ model, labeller=facet_labels)+
    scale_x_discrete(labels=months_first_letter_list)+
    theme_bw() +
    ggtitle(paste("Water Year", water_year))+
    scale_fill_manual(values=crop_palette)+
    scale_y_continuous(labels = axis_units, limits = c(0, 300000)) +
    ylab("Thousand Acre-Feet") +
    theme(#panel.border = element_blank(),
          panel.grid.major = element_blank(),
          plot.title = element_text(hjust = 0.5),
          panel.grid.minor = element_blank(),
          axis.title.x = element_blank(),
          #axis.text.x=element_text(angle=90, hjust=1, vjust=0.5),
          legend.position="bottom", # position of legend or none
          legend.direction="horizontal", # orientation of legend
          legend.title= element_blank(), # no title for legend
          legend.key.size = unit(0.5, "cm"))+ # size of legend
          theme(strip.background = element_blank(), strip.placement = "outside",
                axis.line.x = element_line(color="black", size = 1),
                axis.line.y = element_line(color="black", size = 1))
p

```






# Fieldstation Level Analysis


## Fieldstation vs Remote methods facet plot for Water Year 2015
```{r field_v_remote_2015, echo=FALSE, fig.width = 7, fig.height = 5}

# remove D4
fieldstations_data_2015_filter <- fieldstations_data_2015 %>% filter(!Station %in% c("D4"))

# manually add asterisk to D3 and D4 in the dataframe
fieldstations_data_2015_filter$Station <- gsub('D3', 'D3*', fieldstations_data_2015_filter$Station)
fieldstations_data_2015_filter$Station <- gsub('D4', 'D4*', fieldstations_data_2015_filter$Station)

# ETo should be a horizontal line
eto_df <- fieldstations_data_2015_filter %>% filter(model=="eto") %>% select(Station, ET_mean) %>% rename(eto=ET_mean) %>% distinct(Station, eto)
field_wo_eto <- fieldstations_data_2015_filter %>% filter(!model=="eto")

# join the eto data with the model data
d <- left_join(field_wo_eto, eto_df, by=c("Station"))

# remove EC points
#d <- d %>% filter(!measure_type=='EC')

# spread the measure type to two columns EC and SR
ds <- d %>% spread(measure_type, field_mean_et)


p <- ggplot(ds, aes(x=model, y=ET_mean, color=model))+
  geom_point(size=3.5)+
  geom_hline(aes(yintercept=SR), colour="#990000", linetype="dashed", lwd=1.5, alpha=0.5)+
  geom_hline(aes(yintercept=EC), colour="#990000", linetype="solid", lwd=1.5, alpha=0.5)+
  geom_hline(aes(yintercept=eto), colour="#000000", linetype="dashed", lwd=1.5, alpha=0.5)+
  geom_text(aes(x=6.35,y=eto), size=4, nudge_y = 0.2, colour="#000000", label='ETo')+
  geom_text(aes(x=6.35,y=SR), size=4, nudge_y = 0.2, colour="#990000", label='ETa')+
  #geom_text(aes(x=6.35,y=EC), size=4, nudge_y = -0.2, colour="#990000", label='EC')+
  ylab("Model ET (mm/day)")+
  xlab("Bare Field Station")+
  scale_color_manual(values=model_palette, labels=methods_named_list)+
  theme_bw()+
  theme(panel.border = element_rect(colour = "black", fill="NA", size=1),
        panel.grid.major = element_blank(),
        plot.title = element_text(hjust = 0.5),
        panel.grid.minor = element_blank(),
        legend.position="bottom", # position of legend or none
        legend.direction="horizontal", # orientation of legend
        legend.title= element_blank(), # no title for legend
        legend.key.size = unit(0.5, "cm"),
        axis.text.y = element_text(size=12),
        axis.text.x = element_blank(),
        axis.title = element_text(size=14))+ 
        guides(colour = guide_legend(nrow = 2, keywidth = 3, keyheight = 1),
        fill = guide_legend(keywidth = 1, keyheight = 1),
        linetype=guide_legend(keywidth = 3, keyheight = 1))+
        facet_wrap(~Station, ncol=5, scales="free_x")+theme(strip.background = element_blank(),
        strip.placement = "outside", 
        strip.text.x = element_text(size = 16))

# add footer
p <- p + labs(caption=footer)

# add plot to rmarkdown output
plot(p)

```

## ETRF plots (pixel level calc in EE)

DETAW & DisAlexi provided seperate ETo. The rest of the methods use spatial cimis.

```{r etrf_ee, echo=FALSE, fig.width = 7, fig.height = 4}
# list of crops to generate
crops_2_gen <- c("Alfalfa", "Almonds", "Corn", "Pasture", "Potatoes", "Rice", "Tomatoes", "Vineyards")

# loop through the crops
for(c in crops_2_gen){
  p <- etrf2(dsa_etrf, c)

  # add footer
  p <- p + labs(caption=footer)
  
  # add plot to rmarkdown output
  plot(p)
}
```


# Calculations for numbers in the report

```{r}
# ET totals for non-ag classes 
# for numbers in text above table 3.

# summarize total acre-ft just using non-crops (ie include = "no")
af_noncrops <- dsa_legal_data %>% filter(include == "no", crop_acft > 0) %>%
    dplyr::group_by(model, region, wateryear) %>% 
    dplyr::summarise(sum_af=sum(crop_acft)) %>% filter(region == "dsa")

# remove sims since they don't account for water or native classes + remove eto data
af_noncrops_wo_sims <- af_noncrops %>% filter(!model %in% c("sims", "eto"))

# average of noncrop by year
af_noncrops_avg <- af_noncrops_wo_sims %>% group_by(wateryear) %>% summarize(mean=mean(sum_af))

avg_2015 <- af_noncrops_avg$mean[af_noncrops_avg$wateryear==2015]
avg_2016 <- af_noncrops_avg$mean[af_noncrops_avg$wateryear==2016]

# print results
print(paste("Not all methods are capable of estimating ET for other land uses (i.e. urban, open water, and vegetation), but available estimates for those excluded land classes averaged about", round(avg_2015/1000), "TAF in 2015 and increased to", round(avg_2016/1000) , "TAF in 2016"))

```

```{r}

# average evapotranspiration for september to january as percentage of total annual et

# text after coefficient of variation grids

# summarize total acre-ft just using crops (ie include = "yes")
af_annual <- dsa_legal_data %>% filter(region == "dsa") %>% filter(include == "yes", crop_acft > 0) %>% filter(!model == 'eto') %>% 
    dplyr::group_by(model, region, wateryear) %>% 
    dplyr::summarise(sum_af=sum(crop_acft)) 

af_annual_mean <- af_annual %>% group_by(wateryear) %>% summarize(total=mean(sum_af))


# Between selected months (Sept to Jan)
af_select <- dsa_legal_data %>% filter(region == "dsa") %>% filter(include == "yes", crop_acft > 0) %>% filter(!model == 'eto') %>% 
    filter(month %in% c("OCT", "NOV", "DEC", "JAN"))%>%
    dplyr::group_by(model, region, wateryear) %>% 
    dplyr::summarise(sum_af=sum(crop_acft)) 

af_select_mean <- af_select %>% group_by(wateryear) %>% summarize(select=mean(sum_af))


# join annual and select together using water year
af_both <- merge(af_annual_mean, af_select_mean, by="wateryear") %>% mutate(percent=select/total*100)

# both wateryears are approx 22 or 23 % so let's just take the avearage
af_select_percent <- round(mean(af_both$percent), 1)

print(paste("We employed the coefficient of variation to calculate degree of discrepancy among models by crop are higher from September to January. Evapotranspiration estimates in these months represent about", af_select_percent, "% of the annual ET"))
```


```{r eto, echo=FALSE}


# add date
eto_data <- eto %>% mutate(date=date_from_wy_month(wateryear, month)) # add a date field by combining month and water year

# construct the plot object
p <- ggplot(eto_data, aes(date, mean/10, color=model, group=model)) + 
  geom_line(size=1.25)+
  #geom_line(aes(date, mean/10-stdDev/10, color=model, group=model), size=1.25, alpha=0.6, lty="dashed")+
  #geom_line(aes(date, mean/10+stdDev/10, color=model, group=model), size=1.25, alpha=0.6, lty="dashed")+
  scale_x_date(date_breaks="3 month", date_labels  = "%b")+
  coord_cartesian(ylim=c(0, 8))+
  ggtitle("ETo") +
  ylab("ETo (mm/day)") +
  geom_vline(xintercept = as.numeric(as.Date('2015-10-01')),colour="black", linetype="dashed", size=1.25, alpha=0.5)+
  annotate("text", x=as.Date('2015-04-01') , y=8, label="2015", size=5)+
  annotate("text", x=as.Date('2016-04-01') , y=8, label="2016", size=5)+
  scale_color_manual(labels=c(spatial_cimis_eto="Spatial CIMIS", disalexi_eto="DisALEXI", detaw_eto="DETAW"), values=c(spatial_cimis_eto="#619CFF", disalexi_eto="#00BA38", detaw_eto="#F8766D")) +
  theme_bw() +  # change theme simple with no axis or tick marks
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        plot.title = element_text(hjust = 0.5, size=16),
        panel.grid.minor = element_blank(),
        axis.title.x = element_blank(),
        legend.position="bottom", # position of legend or none
        legend.direction="horizontal", # orientation of legend
        legend.title= element_blank(), # no title for legend
        legend.key.size = unit(0.5, "cm"),
        axis.text.y=element_text(size=14),
        axis.text.x=element_text(size=12),
        axis.title.y=element_text(size=14),
        axis.line = element_line(color="black", size=1)
        )
  

p

```

## Region Level Analysis (North, South, West, Yolo, Central)

```{r regions_taf, echo=FALSE}

# loop through both water years 
for(y in 1:length(water_years)){
  wy <- water_years[y]
  
  p <- barchart_TAFxDeltaregions(delta_regions, wy)
  
  # add footer
  p <- p + labs(caption=footer)
  
  # add plot to rmarkdown output
  plot(p)
}

 
```



```{r regions_taf_table}

water_year <- 2016

af <- delta_regions %>%
    filter(!model=='eto')%>%
    filter(include == "yes", wateryear==water_year, crop_acft > 0) %>%
    dplyr::group_by(model, region) %>% 
    dplyr::summarise(sum_af=sum(crop_acft)) 
  
  
  # sims doesn't include wet herbaceous or semi ag so we'll add a point 
  # that represents the average of all methods for herbaceous/sub irrigated pasture.
  
  # To calculate the estimate, take the total acre-feet for the sims and subtract the values for the herb and semi-ag classes. Then,
  # calculate the mean value from the other models for these two classes and add it to the total.
  
  # total acre-feet for sims for the two classes (to subtract from the total since the EE output has a non-zero value)
  sims_herb_semi <- delta_regions %>% filter(model=="sims", wateryear==water_year, level_2 %in% c(2008, 913)) %>%
    dplyr::group_by(region) %>% 
    dplyr::summarise(sum_wetherb_semiag=sum(crop_acft))
  
  # Calculate the average for herb + semi-ag for the other models for the two regions
  mean_herb_semi <- delta_regions %>% filter(!model=="sims", !model=="eto") %>%
    filter(wateryear==water_year, level_2 %in% c(2008, 913)) %>%
    dplyr::group_by(model, region) %>%
    dplyr::summarise(sum_af=sum(crop_acft)) %>%
    dplyr::group_by(region) %>%
    dplyr::summarise(mean_2classes=mean(sum_af))
  
  # get sims TAF results for DSA
  sims <- af %>% filter(model=="sims")
  
  # join the sims sum for herb_semi and the mean value for herb_semi for the two regions
  sims_joined <- dplyr::left_join(dplyr::left_join(sims, sims_herb_semi, by="region"), mean_herb_semi, by="region")
  
  # subtract the herb_sumi total from the orignal value and add the mean for the two crops
  sims_joined_est <- sims_joined %>% mutate(sum_af_est=(sum_af - sum_wetherb_semiag + mean_2classes))
  
  # clean up columns by droping intermediate columns and renaming estimate to sum_af
  sims_est <- sims_joined_est %>% dplyr::select(c(model, region, sum_af_est)) %>% rename(sum_af=sum_af_est)
  
  # remove sims from af_crop_filter and add in the new estimated results
  af_wo_sims <- af %>% filter(!model=="sims")
  bind <- bind_rows(af_wo_sims,  sims_est)

  
region_avg <- bind %>% group_by(region) %>% summarize(mean(sum_af))

print(region_avg)

```



